---
title: "Statistical Learning in Movies"
author: "Max Sellers, Claire Martino, Ari Augustine, and Rim Nassiri"
output:
  pdf_document: default
  html_document:
    df_print: paged
---

```{r echo=FALSE, warning=FALSE, error=FALSE, message=FALSE}
# if not installed use: install.packages("package_name")
suppressWarnings(library(dplyr))
suppressWarnings(library(readxl))
suppressWarnings(library(psych))
suppressWarnings(suppressPackageStartupMessages(library(ggplot2)))
```

```{r echo=FALSE}
# TODO: change the Actual Time format with just hours and min
movies_data= read.csv("MOVIES.csv")
```


### Abstract
In

## I. Introduction

## II. Methods

### a. Cleaning Data
```{r}
movies_data = movies_data %>%
  group_by(title) %>%
  filter(budget >= 1000000 & revenue >= 1000000 & vote_count > 100) %>%
  arrange(desc(vote_average))
```

### b. Choosing Important Values on Dataset
```{r}

#number of cateogories and varaibles in the dataset
catagories = ncol(movies_data)
var = nrow(movies_data)

#means of variables that will be using in analysis
mean_budget = round(mean(movies_data$budget),2)
mean_popularity = round(mean(movies_data$popularity),2)
mean_revenue = round(mean(movies_data$revenue),2)

```

### c. Training Data

### d. Testing Data

## III. Results

### a. Graphs of Specific Data

### b. Linear Regression Summary and Line
```{r}
#linear regression model of budget vs vote_average
linear_1 = lm(budget ~ vote_average, data = movies_data)
summary_linear1 = summary(linear_1)
summary_linear1

r_2_lin_1 = summary_linear1$r.squared
r_2_lin_1

#graph of above data
plot(budget~ vote_average, data= movies_data, xlab = "Voter Average", ylab = "Budget", main = "Scatterplot of Budget vs Voter Average")
abline(linear_1, col = "red")

#plot(linear_1)

```

Since the Pr(>|t|) value of voter average is <2e-16, and this value is less than the standard level of significance of 0.05, this shows that there is a statistically significant relationship between voter average and budget.

In order to assess the relationship between the predictor and the response variable, you must look at the $R^2$ value. In this case, $R^2$ = `r r_2_lin_1`. Since this value is closer to 0 than it is 1, this indicates a weak relationship between voter average and budget.


```{r}
#linear regression model of revenue vs popularity
linear_2 = lm(revenue ~ popularity, data = movies_data)
summary_linear2 = summary(linear_2)

r_2_lin_2 = summary_linear2$r.squared
r_2_lin_2

#plot(linear_2)

plot(revenue ~ popularity, data= movies_data, xlab = "Popularity", ylab = "Revenue", main = "Scatterplot of Revenue vs Popularity")
abline(linear_2, col = "purple")

linear_3 = lm(vote_average ~ budget, data = movies_data)
summary_linear3 = summary(linear_2)

r_2_lin_3 = summary_linear3$r.squared
r_2_lin_3

#plot(linear_2)

plot(vote_average ~ budget, data= movies_data, xlab = "Popularity", ylab = "Revenue", main = "Scatterplot of Revenue vs Popularity")
abline(linear_3, col = "purple")
```

Since the Pr(>|t|) value of popularity is <2e-16, and this value is less than the standard level of significance of 0.05, this shows that there is a statistically significant relationship between popularity and revenue.

In order to assess the relationship between the predictor and the response variable, you must look at the $R^2$ value. In this case, $R^2$ = `r r_2_lin_2`. Since this value is closer to 0 than it is 1, this indicates a mildly weak relationship between revenue and popularity.

### c. Multiple Linear Regression Summary
```{r}
multi_linear = lm(revenue ~ budget + vote_average, data = movies_data)
summary(multi_linear)

multi_linear2 = lm(vote_average ~ budget + revenue + popularity + runtime, data = movies_data)
summary(multi_linear2)

#confidence interval insert here 
```


## IV. Discussion


